### 2018-02-09 plos-latex-template/manuscript.Rnw
cache chunks for faster writing

\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
Ethomics, a quantitative and high-throughput approach to animal behaviour, is a new and exciting field.
The recent development of automatised methods that can score various behaviours on a large number of animals
% brings the promise of a powerful and original insight
provides biologists with an unprecedented tools to decipher these complex phenotypes. 
Analysing ethomics data comes with many challenges that are, conceptually, independent of the acquisition platform used.
However, there is, little effort in providing a generic framework to specifically analyse multiple and long behavioural time series.
We developed the \texttt{rethomics} framework, a suite of \texttt{R} packages that altogether offer utilities to:
import, store, visualise and analyse behavioural data.
In this article, we describe it and show an example of its application to the blooming field of sleep and circadian rhythm in fruit fly.
The \texttt{rethomics} framework is available and documented at \href{https://rethomics.github.io}{https://rethomics.github.io}.

>>>>

\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
Ethomics, a quantitative and high-throughput approach to animal behaviour, is a new and exciting field.
The recent development of automatised methods that can score various behaviours on a large number of animals
% brings the promise of a powerful and original insight
provides biologists with an unprecedented tools to decipher these complex phenotypes. 
Analysing ethomics data comes with many challenges that are independent of these acquisition platform.
However, there is little effort in providing a generic framework to specifically analyse multiple and long behavioural time series.
We developed the \texttt{rethomics} framework, a suite of \texttt{R} packages that altogether offer utilities to:
import, store, visualise and analyse behavioural data.
In this article, we describe it and show an example of its application to the blooming field of sleep and circadian rhythm in fruit fly.
The \texttt{rethomics} framework is available and documented at \href{https://rethomics.github.io}{https://rethomics.github.io}.

### 2018-02-09 plos-latex-template/results.Rnw
cache chunks for faster writing

<<echo=FALSE, message=FALSE>>=
opts_knit$set(root.dir = "results")
@

@@

<<>>=
metadata <- link_dam2_metadata("metadata.csv",".")    # linking
# print(metadata)                                     # check metadata
dt <- load_dam2(metadata)                             # loading
summary(dt)                                           # quick summary
@

@@

<<message=F, results="hide">>=
# baseline subtraction -- note the use of xmv
dt[,t := t - days(xmv(baseline_days))]
dt[, moving :=  activity > 0]  
summary(dt)
@

@@

<<message=FALSE, rsults="hide">>=
dt[, label := interaction(1:.N, genotype), meta=T]
# print(dt)
@

@@

<<>>=
# make a ggplot object with label on the y and moving on the z axis
fig3A <- ggetho(dt, aes(y=label, z=moving)) +  
  # show data as a tile plot
  # that is z is a pixel whose intensity maps moving
  stat_tile_etho() +                  
  # add layers to draw annotations to show L and D phases
  # as white and black, respectivelly
  # the first layer is for the baseline (until t=0)
  stat_ld_annotations(x_limits = c(dt[,min(t)], 0)) +
  # in the 2nd one, we start at 0 and use grey 
  # instead of black as we work in LL
  stat_ld_annotations(x_limits = c(0, dt[,max(t)]), 
                      ld_colours = c("white", "grey"))
@

Activity of dead or escaped animals is falsely scored as long series of zeros.
Our \texttt{sleepr} package offer a tool to detect and remove such artefactual data.
The updated version can be visualised in Fig~\ref{fig:fig-3}B.
<<>>=
# remove data after death
dt <- sleepr::curate_dead_animals(dt, moving)
# same as above
fig3B <- ggetho(dt, aes(y=label, z=moving)) +  
    stat_tile_etho() +                  
    stat_ld_annotations(x_limits = c(dt[,min(t)], 0)) +
    stat_ld_annotations(x_limits = c(0, dt[,max(t)]), 
                        ld_colours = c("white", "grey"))
@

For the purpose of this example, we keep only individuals that have \emph{at least five days in LL}.
<<>>=
# for each id, we check for validity
valid_dt <- dt[ , .(valid = max(t) > days(5)), by=id]
# a vector of all valid ids
valid_ids <- valid_dt[valid == T, id]
# filter dt with the valid ids
dt <- dt[id %in% valid_ids]
summary(dt)
@

@@

<<>>=
figS1A <- ggetho(dt, aes(z = moving), multiplot = 2) +
            stat_bar_tile_etho() + 
            facet_wrap( ~ label, ncol=4) +
            scale_y_discrete(name="Day")
@

@@

<<>>=
dt_ll <- dt[t > days(1)]
per_dt <- periodogram(moving, 
                    	dt_ll, 
                    	resample_rate = 1/mins(10),
                    	FUN=chi_sq_periodogram)

@@

<<r, eval=TRUE, echo=FALSE, message=FALSE, result="hide">>=
################################ CODE BELOW FOR COSMETIC ALTERATIONS OF FIGURES

@@

fig3 <- cowplot::plot_grid(
                          legend,
                          fig3A +  my_theme,
                          fig3B +  my_theme,
                          nrow=1,
                          labels=c("", "A", "B"),
                          rel_widths = c(1, 2, 2),
                          label_size = 20,
                          axis="b",
                          align="h"
                          )
pdf("../fig/fig-3.pdf",w=12,h=12)
print(fig3)
dev.off()

@@

pdf("../fig/fig-S1.pdf",w=12,h=16)
print(figS1)
dev.off()
@

>>>>

<<echo=FALSE, message=FALSE, cache=TRUE>>=
opts_knit$set(root.dir = "results")
@

@@

<<cache=TRUE>>=
metadata <- link_dam2_metadata("metadata.csv",".")    # linking
# print(metadata)                                     # check metadata
dt <- load_dam2(metadata)                             # loading
summary(dt)                                           # quick summary
@

@@

<<message=F, results="hide", cache=TRUE>>=
# baseline subtraction -- note the use of xmv
dt[,t := t - days(xmv(baseline_days))]
dt[, moving :=  activity > 0]  
summary(dt)
@

@@

<<message=FALSE, results="hide", cache=TRUE>>=
dt[, label := interaction(1:.N, genotype), meta=T]
print(dt)
@

@@

<<cache=TRUE>>=
# make a ggplot object with label on the y and moving on the z axis
fig3A <- ggetho(dt, aes(y=label, z=moving)) +  
  # show data as a tile plot
  # that is z is a pixel whose intensity maps moving
  stat_tile_etho() +                  
  # add layers to draw annotations to show L and D phases
  # as white and black, respectivelly
  # the first layer is for the baseline (until t=0)
  stat_ld_annotations(x_limits = c(dt[,min(t)], 0)) +
  # in the 2nd one, we start at 0 and use grey 
  # instead of black as we work in LL
  stat_ld_annotations(x_limits = c(0, dt[,max(t)]), 
                      ld_colours = c("white", "grey"))
@

Activity of dead or escaped animals is falsely scored as long series of zeros.
Our \texttt{sleepr} package offer a tool to detect and remove such artefactual data.
The updated version can be visualised in Fig~\ref{fig:fig-3}B.

<<cache=TRUE>>=
# remove data after death
dt <- sleepr::curate_dead_animals(dt, moving)
# same as above
fig3B <- ggetho(dt, aes(y=label, z=moving)) +  
    stat_tile_etho() +                  
    stat_ld_annotations(x_limits = c(dt[,min(t)], 0)) +
    stat_ld_annotations(x_limits = c(0, dt[,max(t)]), 
                        ld_colours = c("white", "grey"))
@

For the purpose of this example, we keep only individuals that have \emph{at least five days in LL}.
<<cache=TRUE>>=
# for each id, we check for validity
valid_dt <- dt[ , .(valid = max(t) > days(5)), by=id]
# a vector of all valid ids
valid_ids <- valid_dt[valid == T, id]
# filter dt with the valid ids
dt <- dt[id %in% valid_ids]
summary(dt)
@

@@

<<cache=TRUE>>=
figS1A <- ggetho(dt, aes(z = moving), multiplot = 2) +
            stat_bar_tile_etho() + 
            facet_wrap( ~ label, ncol=4) +
            scale_y_discrete(name="Day")
@

@@

<<cache=TRUE>>=
dt_ll <- dt[t > days(1)]
per_dt <- periodogram(moving, 
                    	dt_ll, 
                    	resample_rate = 1/mins(10),
                    	FUN=chi_sq_periodogram)

@@

<<r, eval=TRUE, echo=FALSE, message=FALSE, result="hide",cache=TRUE>>=
################################ CODE BELOW FOR COSMETIC ALTERATIONS OF FIGURES

@@

fig3 <- cowplot::plot_grid(
                          legend,
                          fig3A +  my_theme,
                          fig3B +  my_theme,
                          nrow=1,
                          labels=c("", "A", "B"),
                          rel_widths = c(1, 2, 2),
                          label_size = 20,
                          axis="b",
                          align="h"
                          )
pdf("../fig/fig-3.pdf",w=12,h=12)
print(fig3)
tmp  <- dev.off()

@@

pdf("../fig/fig-S1.pdf",w=12,h=16)
print(figS1)
tmp <- dev.off()
@
